---
title: "p8105_hw3_ys3508"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)

library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```
# Problem 1
## load package
```{r}
library(p8105.datasets)

```

## Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```
## Answer Questions
### Problem 1.1
Aisles_id
```{r}
instacart %>%
  group_by(aisle) %>%
  summarise(n_aisles = n()) %>%
  arrange(desc(n_aisles))
```

134 aisles are there, and fresh vegetables and fresh fruits are the most items ordered from.

### Problem 1.2
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
bar_aisle <- instacart %>% 
  count(aisle) %>%
  filter(n > 10000) %>% # limiting this to aisles with more than 10000 items ordered
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) +
    geom_bar(stat='identity') +
    xlab("Aisle")+
    ylab("The number of items ordered")+
  ggtitle("The number of items ordered in each aisle (aisle > 10000)") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_aisle

```


### Problem 1.3 
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle == "baking ingredients" | aisle ==  "dog food care" | aisle ==  "packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(product_ranking = min_rank(desc(n))) %>% # rank from the most popular to less polular
  filter(product_ranking < 4) %>%  # list top 3 in each aisle
  arrange(desc(n)) %>%  # order the table by n
  knitr::kable() # making the table
```


### Problem 1.4 
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)


```

# Problem 2
## load data
```{r}
accel = 
  read_csv(
    "data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(weekday = ifelse(day == "Saturday" | day == "Sunday" , "weekend", "weekday"))  %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute", 
    names_prefix = "activity_",
    values_to = "activity_counts") 
accel
```
## Answer Questions
### Probel 1.1
The resulting dataset have 6 variables, including "week", "day_id", "day", "weekday", "minute" and "activity_counts", with 50400 observations in total.

### Problem 1.2

# Problem 3
## load data
```{r}
data("ny_noaa")
```

