---
title: "p8105_hw3_ys3508"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)

library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```
# Problem 1
## load package
```{r}
library(p8105.datasets)

```

## Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```
## Answer Questions
### Problem 1.1
Aisles_id
```{r}
instacart %>%
  group_by(aisle) %>%
  summarise(n_aisles = n()) %>%
  arrange(desc(n_aisles))
```

134 aisles are there, and fresh vegetables and fresh fruits are the most items ordered from.

### Problem 1.2
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
bar_aisle <- instacart %>% 
  count(aisle) %>%
  filter(n > 10000) %>% # limiting this to aisles with more than 10000 items ordered
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) +
    geom_bar(stat='identity') +
    xlab("Aisle")+
    ylab("The number of items ordered")+
  ggtitle("The number of items ordered in each aisle (aisle > 10000)") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_aisle

```


### Problem 1.3 
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle == "baking ingredients" | aisle ==  "dog food care" | aisle ==  "packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(product_ranking = min_rank(desc(n))) %>% # rank from the most popular to less polular
  filter(product_ranking < 4) %>%  # list top 3 in each aisle
  arrange(desc(n)) %>%  # order the table by n
  knitr::kable() # making the table
```


### Problem 1.4 
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)


```

# Problem 2
## load data
```{r}
accel = 
  read_csv(
    "data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute", 
    names_prefix = "activity_",
    values_to = "activity_counts") %>% 
  mutate(
    minute = as.numeric(minute),
    activity_counts = as.numeric(activity_counts),
    week = as.factor(week),
    day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")),
    weekday_vs_weekend = ifelse(day == "Saturday" | day == "Sunday" , "weekend", "weekday")) %>%
  arrange(day)

```
## Answer Questions
### Probelm 1.1
The resulting dataset have 6 variables, including "week", "day_id", "day", "weekday_vs_weekend", "minute" and "activity_counts", with 50400 observations in total.

### Problem 1.2
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}

accel_2 <- accel %>%
  group_by(week, day) %>%
  summarise(total_activity = sum(activity_counts)) %>%
  pivot_wider(
    names_from = "day",
    values_from = "total_activity"
  ) %>%
  relocate(week, Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday) 
accel_2

```

I didn't find regular trends. However, one thing worth noticing is that on weeks 4 and 5's Saturdays, his activity counts are both low. 

### Problem 1.3
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel %>%
  group_by(day, minute) %>% 
  summarize(avg_counts = mean(activity_counts)) %>%
  ggplot(aes(x = minute, y = avg_counts, color = day)) +
  geom_point()+
  geom_smooth(methd = lowess, span = .3) +
  scale_x_continuous(
    breaks = c(0:24) * 60,
    labels = c(0:24)
  ) +
  labs(
    title = "24-hour activity time courses for each day",
    x = "Hours for each day",
    y = "Average activity counts")+
    theme(legend.position = "right", axis.text.x=element_text(angle=45, hjust=.5, vjust=0.5)) 

```
According to this graph, this patient's activity level is typically low at 11 p.m. to 5 a.m. His most active activity peak is around 21 o'clock on Friday, and his second most active is around 10 o'clock on Sunday morning.

# Problem 3
## load data
```{r}
data("ny_noaa")
```

