---
title: "Data Tidying"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

## load package
```{r}
library(tidyverse) 
library(readxl)
```


## Problem 1
### Load/Clean Data
```{r}
nyc_mta = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv", 
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) %>% 
  rename(ADA_compliance = ada)%>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))

```

These data are not "tidy" as they now stand; route number and route should both be variables. That is, we would need to change "route" variables from wide to long format in order to have a clean dataset. 
### Answer Questions
```{r}
nyc_mta %>% 
  select(station_name, line) %>% 
  distinct

nyc_mta %>% 
  filter(ADA_compliance == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

There are 465 distinct stations, and 84 of them are ADA compliant.

```{r}
nyc_mta %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

```{r}
nyc_mta %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

nyc_mta %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ADA_compliance == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

## Problem 2
### Load/Clean Data

```{r}
mr_trash_wheel = 
  readxl::read_excel("./data/Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel", range = "A2:N550") %>%
    janitor::clean_names() %>%
  filter(dumpster >= 1) %>% 
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(trash_wheel_source = "Mr.")

mr_trash_wheel
```

```{r}
pro_trash_wheel = 
  readxl::read_excel("./data/Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel", range = "A2:M97") %>%
    janitor::clean_names() %>%
  filter(dumpster >= 1) %>%
  mutate(trash_wheel_source = "Professor") %>% 
  mutate(year = as.character(year))

pro_trash_wheel
```

### Join Two Dataset
```{r}
trash_wheel_data = 
  full_join(mr_trash_wheel, pro_trash_wheel, by  = c('dumpster','month','year','date','weight_tons','volume_cubic_yards','plastic_bottles','polystyrene','cigarette_butts','glass_bottles', 'grocery_bags', 'chip_bags', 'homes_powered', 'trash_wheel_source'))

trash_wheel_data

```

### Answer Questions

In the resulting dataset, there are 641 observations. As an example, the variable "dumpster" has 641 641 observations, of which 547 are from the "Mr. Trash Wheel" dataset and the remaining 94 from the "Professor Trash Wheel" dataset.

```{r}
trash_wheel_data %>% 
  select(dumpster, trash_wheel_source) %>% 
  distinct()

trash_wheel_data %>% 
  filter(trash_wheel_source == "Mr.") %>% 
  select(dumpster, trash_wheel_source) %>% 
  distinct()

trash_wheel_data %>% 
  filter(trash_wheel_source == "Professor") %>% 
  select(dumpster, trash_wheel_source) %>% 
  distinct()
```
The total weight of trash collected by Professor Trash Wheel was  190.12 tons.

```{r}
trash_wheel_data %>% 
  filter(trash_wheel_source == "Professor") %>% 
  summarise(total_weight = sum(weight_tons, na.rm = TRUE))

```

The total number of sports balls collected by Mr. Trash Wheel in 2020 was 856.

```{r}
trash_wheel_data %>% 
  filter(trash_wheel_source == "Mr." & year == "2020") %>% 
  summarise(total_sports_balls = sum(sports_balls, na.rm = TRUE))

```

## Problem 3
### Load/Clean Data

```{r}
pols = 
  read_csv("/Users/ovo/Desktop/FALL/Data Science/Assignments/p8105_hw2_ys3508/data/p8105_hw2_ajg2202/data/fivethirtyeight_datasets/pols_month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c('year', 'month', 'day'), sep = '-') %>% 
  mutate(month = ifelse(month == "01", "jan", 
                        ifelse(month == "02", "feb", 
                               ifelse(month == "03", "mar", 
                                      ifelse(month == "04", "apr", 
                                             ifelse(month == "05", "may", 
                                                    ifelse(month == "06", "jun", 
                                                           ifelse(month == "07", "jul", 
                                                                  ifelse(month == "08", "aug",
                                                                         ifelse(month == "09", "sep", 
                                                                                ifelse(month == "10", "oct", 
                                                                                       ifelse(month == "11", "nov", "dec")))))))))))) %>%  
  mutate(president = ifelse(prez_dem == "1", "Democratic", "Republican")) %>% 
  select(-prez_dem, -prez_gop, -day)

pols
```

```{r}
snp = 
  read_csv("/Users/ovo/Desktop/FALL/Data Science/Assignments/p8105_hw2_ys3508/data/p8105_hw2_ajg2202/data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(col = date, into = c('month', 'day', 'year'), sep = '/') %>% 
  mutate(month = ifelse(month == "01", "jan", 
                        ifelse(month == "02", "feb", 
                               ifelse(month == "03", "mar", 
                                      ifelse(month == "04", "apr", 
                                             ifelse(month == "05", "may", 
                                                    ifelse(month == "06", "jun", 
                                                           ifelse(month == "07", "jul", 
                                                                  ifelse(month == "08", "aug",
                                                                         ifelse(month == "09", "sep", 
                                                                                ifelse(month == "10", "oct", 
                                                                                       ifelse(month == "11", "nov", "dec")))))))))))) %>% 
  mutate(year = as.numeric(year)) %>% 
  mutate(year = ifelse(year > 90, year  + 1900, year  + 2000)) %>%
  mutate(year = as.character(year)) %>% 
  select( -day) %>% 
  select("year", "month", everything())
 

snp
```


```{r}
unemployment = 
  read_csv("/Users/ovo/Desktop/FALL/Data Science/Assignments/p8105_hw2_ys3508/data/p8105_hw2_ajg2202/data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment") %>%
  mutate(year = as.character(year)) %>% 
  select("year", "month", everything())
 

unemployment
```

### merge dataset
```{r}
# merging snp into pols

snp_pols = merge(pols, snp, by=c("year","month"),all.x=TRUE)
snp_pols
# merging unemployment into the result
total = left_join(snp_pols, unemployment, by=c("year","month"))
total
```

Snp and pols are being merged in the dataset 'snp pols'. This "snp pols" has 10 variables and 1038 observations. The years 1947 through 2015 are included. Important variables include "month" and "year," as well as "gov gop," "sen gop," "rep gop," "gov dem," "sen dem," "rep dem," "president," and "close."

We obtained the dataset "total" with 1038 observations and 11 variables after combining unemployment into the outcome. also ranges from 1947 to 2015. Prior variables included the variable "unemployment."


