---
title: "p8105_hw3_ys3508"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)

library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```
# Problem 1
## load package
```{r}
library(p8105.datasets)

```

## Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```
## Answer Questions
### Problem 1.1
Aisles_id
```{r}
instacart %>%
  group_by(aisle) %>%
  summarise(n_aisles = n()) %>%
  arrange(desc(n_aisles))
```

134 aisles are there, and fresh vegetables and fresh fruits are the most items ordered from.

### Problem 1.2
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
bar_aisle <- instacart %>% 
  count(aisle) %>%
  filter(n > 10000) %>% # limiting this to aisles with more than 10000 items ordered
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) +
    geom_bar(stat='identity') +
    xlab("Aisle")+
    ylab("The number of items ordered")+
  ggtitle("The number of items ordered in each aisle (aisle > 10000)") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
bar_aisle

```


### Problem 1.3 
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle == "baking ingredients" | aisle ==  "dog food care" | aisle ==  "packaged vegetables fruits") %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(product_ranking = min_rank(desc(n))) %>% # rank from the most popular to less polular
  filter(product_ranking < 4) %>%  # list top 3 in each aisle
  arrange(desc(n)) %>%  # order the table by n
  knitr::kable() # making the table
```


### Problem 1.4 
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)


```

# Problem 2
## load data
```{r}
accel = 
  read_csv(
    "data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute", 
    names_prefix = "activity_",
    values_to = "activity_counts") %>% 
  mutate(
    minute = as.numeric(minute),
    activity_counts = as.numeric(activity_counts),
    week = as.factor(week),
    day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")),
    weekday_vs_weekend = ifelse(day == "Saturday" | day == "Sunday" , "weekend", "weekday")) %>%
  arrange(day)

```
## Answer Questions
### Probelm 1.1
The resulting dataset have 6 variables, including "week", "day_id", "day", "weekday_vs_weekend", "minute" and "activity_counts", with 50400 observations in total.

### Problem 1.2
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}

accel_2 <- accel %>%
  group_by(week, day) %>%
  summarise(total_activity = sum(activity_counts)) %>%
  pivot_wider(
    names_from = "day",
    values_from = "total_activity"
  ) %>%
  relocate(week, Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday) 
accel_2

```

I didn't find regular trends. However, one thing worth noticing is that on weeks 4 and 5's Saturdays, his activity counts are both low. 

### Problem 1.3
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel %>%
  group_by(day, minute) %>% 
  summarize(avg_counts = mean(activity_counts)) %>%
  ggplot(aes(x = minute, y = avg_counts, color = day)) +
  geom_point()+
  geom_smooth(methd = lowess, span = .3) +
  scale_x_continuous(
    breaks = c(0:24) * 60,
    labels = c(0:24)
  ) +
  labs(
    title = "24-hour activity time courses for each day",
    x = "Hours for each day",
    y = "Average activity counts")+
    theme(legend.position = "right", axis.text.x=element_text(angle=45, hjust=.5, vjust=0.5)) 

```
According to this graph, this patient's activity level is typically low at 11 p.m. to 5 a.m. His most active activity peak is around 21 o'clock on Friday, and his second most active is around 10 o'clock on Sunday morning.

# Problem 3
## load data
```{r}
data("ny_noaa")

```

##Answer Questions
### Problem 3.1 
A short description of the dataset: 
```{r}
ny_noaa
```

The raw dataset `ny_noaa` contains 2,595,176 `r nrow(ny_noaa)` observations with 7 `r ncol(ny_noaa)` variables, which include `id`, `date`, `prcp`, `snow`, `snwd`, `tmax` and `tmin`. 
The following are important variables: id` means weather station ID, while `date` represents date of observation. `prcp` means precipitation (tenths of mm), `snow` means snowfall (mm) and `snwd` means snow depth (mm). `tmax` and `tmin` each represent the maximum temperature and minimum temperature(tenths of degrees C) for the day. 
Missing data appear in 5 variables: 'prcp', 'snow' , 'snwd',  'tmax' and '
tmin'. The missing data in the 'prcp', 'snow' , 'snwd' is be an issue.

### Problem 3.2
Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
```{r}
ny_noaa <- ny_noaa %>%
  janitor::clean_names() %>% 
  drop_na(prcp, snow, snwd, tmax, tmin) %>% #drop na in the dataset
  separate(
    date, 
    into = c("year", "month", "day"), 
    sep = "-", 
    convert = TRUE) %>% 
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    prcp = as.numeric(prcp)) %>% 
  mutate(
    tmax = tmax / 10,
    tmin = tmin / 10,
    prcp = prcp / 10) #report the data as Celsius;

  
ny_noaa %>%
  group_by(snow) %>% 
  summarise(n_snow = n()) %>%
  arrange(desc(n_snow))
```

After removing all the NAs, the most commonly observed values in snow are 0, 25, and 13. Because in most time of the year, there is no snow, 0 is reasonable.

### Problem 3.3
Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
avgtmax <- ny_noaa %>%
  filter(month == "1" | month == "7" ) %>%
  group_by(id, year, month) %>%
  summarise(avg_tmax = mean(tmax, na.rm = TRUE)) %>% 
  ggplot(aes(x = year, y = avg_tmax, color = month)) + 
  geom_point() +
  geom_smooth(methd = lowess, span = .3) +
  facet_grid(~ month) +
  labs(
  title = "The average max temperature in January and in July in each station across years",
  x = "Year",
  y ="The average max temperature", 
  labels = scales::comma) +
    theme(axis.text.x=element_text(angle=45, hjust=.5, vjust=0.5)) 
avgtmax
```

The plot shows that the average temperature in July ranges between -20 and 10 degrees Celsius and in July ranges between 20 and 30 degrees Celsius, implying that the average maximum temperature in July is significantly higher than it is in January. The lines in this image are roughly parallel. However, there are some outliers, such as the January 1982 point and the July 1988 point.